### 2019-12-26 paper2/article.tex
Added paper 7.

%--%\documentclass[aps,pre,twocolumn,amsmath,amssymb,amsfonts]{revtex4-1}
%--%%\documentclass[aps,pre,twocolumn,amsmath,amssymb,amsfonts]{revtex4}
%--%\usepackage{epsfig}
%--%\usepackage{graphicx}
%--%\usepackage{dcolumn}
%--%\usepackage{bm}\usepackage[titletoc]{appendix}
%--%
%--%
%--%\usepackage{graphicx}
%--%\usepackage{subfigure}
%--%\usepackage{amsfonts}
%--%\usepackage{amsgen}
%--%\usepackage{amsbsy}
%--%%\usepackage{amsmath}
%--%%\usepackage{amssymb}
%--%%\usepackage{mathrsfs}
%--%%\usepackage{mathtools}
%--%\usepackage{epstopdf}
%--%\usepackage{caption}
%--%\usepackage{tabularx}
%--%\usepackage{tabu}
%--%\usepackage{xcolor}
%--%\usepackage{nicefrac}
%--%\usepackage{enumitem}
%--%\usepackage[normalem]{ulem} %To strikeout text
%--%
%--%
%--%\newcommand\Ge{\textcolor{blue}{Ge:}}
%--%\newcommand{\e}[1]{\ensuremath{\times 10^{#1}}}
%--%
%--%\newcommand{\red}[1]{{\color{red}#1}}
%--%\newcommand{\blue}[1]{{\color{blue}#1}}
%--%
%--%
%--%\begin{document}
%--%
%--%\author{Itzhak Fouxon$^{1,2}$} \email{itzhak8@gmail.com}  \author{Zhouyang Ge$^3$}\email{zhoge@mech.kth.se} \author{Luca Brandt$^3$}\email{luca@mech.kth.se} \author{Alexander Leshansky$^1$}\email{lisha@tx.technion.ac.il}
%--%\affiliation{$^1$ Department of Chemical Engineering, Technion, Haifa 32000, Israel}
%--%\affiliation{$^2$ Department of Computational Science and Engineering, Yonsei University, Seoul 120-749, South Korea}
%--%\affiliation{$^3$ Linn\'{e} FLOW Centre and SeRC (Swedish e-Science Research Centre), KTH Mechanics, SE-100 44 Stockholm, Sweden}
%--%
%--%\begin{abstract}
%--%
%--%We construct a boundary integral representation for the low-Reynolds-number flow in a channel in the presence of freely-suspended particles (or droplets) of arbitrary size and shape. We demonstrate that lubrication theory holds away from the particles at horizontal distances exceeding the channel height and derive a multipole expansion of the flow which is dipolar to the leading approximation. We show that the dipole moment of an arbitrary particle is a weighted integral of the stress and the flow at the particle surface, which can be determined numerically. We introduce the equation of motion that describes hydrodynamic interactions between arbitrary, possibly different, distant particles, with interactions determined by the product of the mobility matrix and the dipole moment. Further, the problem of three identical interacting spheres initially aligned in the streamwise direction is considered and the experimentally observed ``pair exchange" phenomenon is derived analytically and confirmed numerically. For non-aligned particles, we demonstrate the formation of a configuration with one particle separating from a stable pair. Our results suggest that in a dilute initially homogenous particulate suspension flowing in a channel the particles will eventually separate into singlets and pairs.
%--%
%--%
%--%
%--%\end{abstract}
%--%
%--%\title{Integral representation of channel flow with interacting particles}
%--%\maketitle

Hydrodynamic interactions among particles flowing in the fluid confined between two parallel walls at low Reynolds number have recently attracted a considerable attention \cite{carba,d1,d2,d3,d4,d5,d6,d7,anom,7,tlusty,tl2006,tl2012,is,sb,f1,f2,f3,f4,f5,f7,tl2014,tabeling,tab0}. The case of particles driven by thermal noise in the absence of a macroscopic flow was studied in \cite{carba,d1,d2,d3,d4,d5,d6,d7,anom,7,tlusty}. The hydrodynamic interactions cause long-range correlations 
in their diffusive motions 
that are measurable even at distances ten times larger than the particle size \cite{carba,anom}. In the case of
%LB macroscopic
pressure-driven Poiseuille or shear flow the particles are, in addition, dragged by the flow \cite{f1,f2,f3,f4,f5,f7,tl2006,tl2012,tl2014,tabeling,tab0,is}.

Identical particles at similar positions inside the channel move at the same velocity if not for hydrodynamic interactions. These interactions induce particle relative motions, which can result in considerable changes of their configuration inside the channel. 
In the case of a large number of particles the interactions cause also chaotic collisions among the particles \cite{tl2014}.

Theoretical progress has mainly relied on the observation that the far flow caused by a particle confined in a channel is a dipolar flow decaying quadratically with the distance \cite{anom}. For disk-like particles with thickness close to the channel width $h$, the dipolar flow and its moments were derived from lubrication theory in \cite{tlusty}. The dipolar flow holds at distances much larger than the disk radius, where it gives also the leading order hydrodynamic interactions among particles \cite{tlusty,tl2006,tl2012,tl2014}.

It was observed in \cite{tl2012}, however, that hydrodynamic interactions of pancake-like disks can also be described at much smaller distances between the disks where dipolar approximation breaks down, yet lubrication theory still holds \cite{Batchelor,lubr,szeri,bruce}. This theory predicts that at distances from the particle boundary much larger than $h$ the depth-averaged flow is an ideal two dimensional flow with potential obeying the Laplace equation. The boundary condition (b.\ c.), derived somewhat heuristically, is the usual ideal flow b.\ c.\ prescribing  the velocity component normal to the particle surfaces \cite{ll}, which allowed to find the hydrodynamic interactions of two close disks, see \cite{tl2012}. Moreover, it was observed that the non-rigidity of the particles makes the lubrication theory valid up to distances from the particles smaller than $h$. The calculation of the hydrodynamic interactions for disks of different radii requires solving the Laplace equation with the help of bipolar coordinates, see \cite{is}.

Recently, a practical applications of hydrodynamic interactions among particles in a channel has been proposed. In particular, it is suggested that the combined action of adhesive (non-hydrodynamic) forces and hydrodynamic interactions between microdroplets can result in the formation of regular particle clusters and can thus be potentially used for the production of new materials \cite{tab0,tabeling}. The hydrodynamic forces are believed to be a significant factor in these structure formation. Though the particles 
forming the structure are in a close proximity in the experiments mentioned above, the hydrodynamic interactions are described phenomenologically by a dipolar flow, formally only valid at larger distances. Despite the use of the far-field dipolar flow beyond its domain of validity, the numerical simulations  in \cite{tab0} showed very good agreement with the experimental results \cite{tabeling}. The above motivates the need for the detailed theoretical study of hydrodynamic interactions among particles in narrow channels.

In this work, we introduce a boundary integral representation of the channel flow in the presence of freely suspended particles. The particles can be rigid or soft (droplets).
The representation does not depend on the particle equation of motion, defined by inertia.
Boundary integral representations are known to be useful in unconfined flows and can also be applied to confined geometries  \cite{hb,ps}.
The flow is here expressed as the sum of the undisturbed Poiseuille flow and an integral over the surfaces of all particles, where the particles can have arbitrary shapes.
The derivation is performed for a pressure-driven flow, but identical considerations can be applied to shear flows.

Our representation results in a formula for calculation of the dipole moment, which was previously available only for the case of disk-like particles. The moment is given in terms of a weighted integral of the
stress tensor and the  flow over the surface(s) of the particle(s). Once this integral is numerically tabulated, the result can be used to approximate the flow in different configurations. Here, we perform simulations for the case of neutrally buoyant rigid spherical particles and compute the integral for different positions of the particle center and different ratios of the particle radius to the channel height, i.e.\ different confinements.

We use this new integral representation to show that the lubrication theory holds at the particle near proximity, closer than what typically expected. As an example, we solve the problem of three aligned particles moving along the line defined by their centers and  the case of three nonaligned particles.
We conjecture that this solution is the attractor to which the long-time evolution of arbitrary initial condition converges. We conclude by proposing a mean field description of strong hydrodynamic interactions of close
particles in a dense suspension.

\section{Integral representation for channel flow with particles}

In this Section we derive the boundary integral representation for channel flow in the presence of an arbitrary number of particles of arbitrary shape (see Fig.\ \ref{fig: setup} where spherical particles are shown for illustration). It is assumed that the Reynolds number is low and the Stokes equations hold. The
derivation uses the reciprocal theorem with the reciprocal flow given by the Stokeslet in a channel \cite{LironMochon}, similarly to the derivations in infinite space, see e. g. \cite{ps}.
In this Section we make no assumptions on the form of the equation of motion of the particles which may change according to the relevance of inertia. The particles can be rigid, droplets or, e.g.\ viscoelastic.

\begin{figure}[t!]
 \begin{center}
\includegraphics[width=.9\columnwidth]{channel_sphere3.pdf}
\end{center}
  \caption{Schematic configuration of spherical particles flowing in the pressure-driven (Poiseuille) flow in a channel.}
 \label{fig: setup}
\end{figure}

The undisturbed flow $\bm u^0$, in the absence of  particles, is the Poiseuille flow driven by the constant pressure gradient $\nabla p^0$,
\begin{eqnarray}&&
u^0_x=\frac{z(z-h)\nabla_x p^0}{2\eta},\ \ \nabla p^0=\eta\nabla^2\bm u^0.\label{ps}
\end{eqnarray}
where $\eta$ is the fluid viscosity, $z$ is the vertical coordinate and $h$ is the channel width. The flow is in $x-$direction, $\nabla_x p^0=-|\nabla_x p^0| \bm{\hat x}$. In the presence of a freely suspended particle we look for the solution of,
\begin{eqnarray}&&%\!\!\!\!\!\!\!\!\!\!\!\!\!
\nabla p=\eta\nabla^2\bm u,\ \ \nabla\cdot\bm u=0,\ \ \ \ \bm u(z=0)=\bm u(z=h)=0,\nonumber\\&&%\!\!\!\!\!\!\!\!\!\!\!\!\!
u_x(\infty)=\frac{z(z-h)\nabla_x p^0}{2\eta},
\label{ds}
\end{eqnarray}
which holds outside the particle. The flow is completely determined when solving for the particle motion, i.e.\ knowing the instantaneous particle position as it determines the boundary condition $\bm u_{S}(\bm x)$ on the 
 particle surface $S$. In the case of a rigid particle, $\bm u_{S}(\bm x)=\bm v+\bm \omega\times (\bm x-\bm y)$ where $\bm v$ and $
\bm \omega$ are the particle translational and angular velocities  and $\bm y=(x_p, y_p, z_p)$ is the coordinate of the particle center of mass. The velocities $\bm v$ and $\bm \omega$ are determined by the solution of the equation for the particle motion coupled with the flow. These velocities could be time-independent as in the case of the steady motion of a neutrally buoyant rigid particle or the case of a non-neutrally buoyant particle after sedimentation when reaching the bottom wall (the theoretical determination of these velocities is impossible generally because of the interaction with the walls). These velocities can also be time-dependent as in the case of
a transient flow  or the gravitational settling of a non-neutrally buoyant particle. If several particles are considered, a time-dependent configuration can be induced by their interactions. We assume here that the time variations are not fast so that the unsteady time-derivative term of the Navier-Stokes equations is negligible (for the steady motion of one particle the time-derivative is the spatial derivative of the flow along the streamline which is small  because of the smallness of the Reynolds number). In the case of droplets the boundary condition on the surface is determined by matching with the inner flow. However there is no need for solving for this inner flow since the detailed form of $\bm u_{S}(\bm x)$ is irrelevant for the derivation of the present representation. The generalization of the problem to the case of many particles is obvious.

{\it Implications of lubrication theory.}---Some conclusions on the flow at distances from the particle much larger than the channel width $h$ can be obtained from lubrication theory \cite{Batchelor,lubr,szeri,bruce}. The lubrication theory predicts that at these distances,
\begin{eqnarray}&&
\bm u=\frac{z(z-h)\nabla p}{2\eta},\label{lubr}
\end{eqnarray}
with a certain $z-$independent $p$. Clearly, at large distances, $p\approx p^0$ at the leading order. The depth-averaged velocity $\bm u_d$ is the ideal potential two-dimensional flow,
\begin{eqnarray}&&
\bm u_d=\nabla\phi,\ \ \phi=-\frac{h^2 p}{12\eta},\ \ \nabla^2\phi=0.\label{dlp0}
\end{eqnarray}
In some cases this helps determining the flow completely.

{\it Large disks.}---Large non-wetting droplets squeezed between the walls of a Hele-Shaw cell have pancake-like shapes. These can be modeled as disks with radius $a\gg h$ and height close to $h$, i.e., they almost fill the entire channel height \cite{tlusty,tl2006,tl2012,tl2014}. In this case, one can use Eqs.~(\ref{lubr})-(\ref{dlp0}) at distances from the body that are much larger than $h$ but much smaller than $a$. Thus the ideal flow holds outside the narrow boundary layer near the particle surface whose characteristic size $l_0$ is of the order $h$, much smaller than the particle horizontal size $a$. We call the layer containing the flow vorticity the viscous layer and assume that there is no flux of mass through the surface of the droplet, which keeps its shape and volume. Thus, in the frame of reference moving with the disk, the normal 
velocity component is zero at the outer boundary of the layer and the ideal flow outside the layer is determined uniquely by this boundary condition. However the geometry of the layer is not always known and the complete determination of the flow not possible. If we are only interested in the flow outside the immediate $l_0-$vicinity of the viscous layer,  we can set the boundary condition for the ideal flow on the disk itself, exploiting the fact that $l_0\ll a$. The flow is then found as the dipole potential \cite{tlusty},
\begin{eqnarray}&&
\phi=-\frac{h^2 p^0}{12\eta}-\frac{\bm d\cdot{\hat r}}{r} ,\ \ \nabla^2\phi=0,\ \ \bm d=a^2\bm v,\label{dl}
\end{eqnarray}
where $\bm d$ is the dipole moment, $\bm v=-v\bm{\hat x}$ is the difference between the disk velocity and $-h^2\nabla p_0/(12\eta)$. It is readily seen that the normal, radial component of the velocity $\nabla \phi$ on the surface of the disk is $[\bm v-h^2\nabla p^0/(12\eta)]\cdot{\hat r}$. Note that the velocity $v>0$ since the particle moves slower than the fluid.

The tangential velocity component on the outer boundary of the viscous layer obtained from Eq.~(\ref{dl}) does not match the tangential velocity of the droplet surface. In contrast with the normal component, which can be considered almost constant through the viscous layer, the tangential component changes quickly through this layer to match the inner flow at the droplet surface. As example, in the limit of high droplet viscosity, the condition on the disk surface is that the flow is the appropriate superposition of translation and solid body rotation. Similar viscous layers occur for rigid bodies oscillating in the fluid \cite{ll}. The ideal flow was obtained in \cite{tlusty} and here we describe how this ideal flow fits the complete equations for the viscous flow.
%We find,
%\begin{eqnarray}&&
%\bm u_d=\frac{2{\hat r}(\bm d\cdot{\hat r})-\bm d}{r^2},\ \ \bm u_d\cdot {\hat r}=\frac{(\bm d\cdot{\hat r})}{a^2}
%\\&&
%(u_d)_x=\nabla \frac{a^2 v x}{r^2}=\frac{a^2 v(y^2-x^2)}{r^4}.
%\end{eqnarray}

{\it Boxes.}---Another case where the flow can be fixed without detailed calculations is the case of a box whose smallest dimension is close to $h$ and the longer dimensions are much larger than $h$. If the box is located in the channel so that the flow is
perpendicular to its longer axis with length $l\gg h$ then far from the ends of the box we find the ideal two-dimensional flow with constant velocity on the line.
The solution for the ideal flow with constant velocity on an infinite linear boundary is the uniform flow. We thus conclude that in the frame moving with the box there is a region of stagnant flow behind the box, whose size is of order $l$.

Generally, the flow can be inferred from the lubrication theory in quite a detail for particles whose horizontal dimensions are much larger than $h$ and whose vertical dimension is close to $h$. The flow outside the narrow viscous layer near the particle surface is $z(z-h)$ times the two-dimensional ideal flow determined by the boundary condition of zero normal velocity on the particle surface. The tangential velocity changes fast across the viscous layer. However, if the particle horizontal dimensions are not large or the dimensions are large but the vertical dimension is not close to $h$, a different approach is needed.

{\it Boundary integral representation from the reciprocal theorem.}---We use the reciprocal theorem \cite{hb,ps} using as the reciprocal flow the solution of \cite{LironMochon} for the point-force or Stokeslet between two parallel plates, i.e.\
\begin{eqnarray}&&
-\nabla p^S+\eta \nabla^2 \bm u^S+\bm g\delta(\bm x-\bm x_0)=0,\ \ \nabla\cdot\bm u^S=0, \label{stokes}\\&& \bm u^S(z=0)=\bm u^S(z=h)=0,\ \ \bm u^S(x^2+y^2\to\infty)=0.\nonumber \end{eqnarray}
Analogously to the flow due to a point-force acting on a viscous fluid in infinite space (e.g. \cite{ps}), the solution depends linearly on the source forcing $\bm g$,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!
\bm u^S(\bm x)=\frac{1}{8\pi \eta}S_{ik}(\bm x, \bm x_0)g_k, \label{vl}
\end{eqnarray}
where we introduced the tensor $S_{ik}$ independent of $\bm g$. Similarly for the stress tensor of the Stokeslet solution we can write,
\begin{eqnarray}&&
\sigma^S_{ik}(\bm x)\!=-p^S\delta_{ik}+\eta\left(\nabla_k u^S_i+\nabla_i u^S_k\right)\!=\!\frac{T_{ilk}(\bm x, \bm x_0)g_l}{8\pi},\label{st}\\&& \nabla_kT_{ilk}\!=\!-8\pi\delta_{il}\delta(\bm x\!-\!\bm x_0),\ \ T_{ilk}\!\!=\!\!-p_l \delta_{ik}\!+\!\nabla_k S_{il}\!+\!\nabla_i S_{kl},\nonumber
\end{eqnarray}
where $T_{ilk}$ is a third-rank tensor independent of $\bm g$ and we defined the $\bm g-$independent $p_l$ by $p^S=\bm p\cdot\bm g/(8\pi)$.
%\begin{eqnarray}&&%\!\!\!\!\!\!\!\!\!\!\!\!\!\!
%T_{ilk}\!\!=-p_l \delta_{ik}\!+\!\nabla_k S_{il}\!+\!\nabla_i S_{kl},\label{strt}
%\end{eqnarray}
%where $p_l$ is independent of $\bm g$.
We use the Lorentz identity,
\begin{eqnarray}&&
\nabla_k\left[u_{i}\sigma^S_{ik}-u^S_i\sigma_{ik}\right]+\bm u\cdot \bm g\delta(\bm x-\bm x_0)=0,\label{lo}
%\\&&\sigma_{ik}=-p\delta_{ik}+\eta\left(\nabla_iu_k+\nabla_ku_i\right).\nonumber
\end{eqnarray}
readily inferred from the Stokes equations with $\sigma_{ik}$ the stress tensor of the flow defined by Eqs.~(\ref{ds}),
\begin{eqnarray}&&
\sigma_{ik}=-p\delta_{ik}+\eta\left(\nabla_iu_k+\nabla_ku_i\right).
\end{eqnarray}
Substituting $\bm u^S$ and $\bm \sigma^S$ from Eqs.~(\ref{vl})-(\ref{st}) in Eq.~(\ref{lo}) and using the above identity we find that,
\begin{eqnarray}&&
8\pi \eta u_{l}(\bm x')\delta(\bm x'-\bm x)=\frac{\partial}{\partial x_k'}\left[S_{il}(\bm x', \bm x)\sigma_{ik}(\bm x')
\right.\nonumber\\&&\left.-\eta u_{i}(\bm x')T_{ilk}(\bm x', \bm x)\right].\label{fd}
\end{eqnarray}
Integrating this equation over $\bm x'$ outside the particles,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
u_l\!=\!f_l\!-\!\sum_n\int_{S_n}\!\!\!\frac{S_{il}(\bm x', \bm x)\sigma'_{ik}dS_k'}{8\pi \eta}\nonumber\\&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
+\sum_n\int_{S_n}\frac{u_{i}(\bm x')T_{ilk}(\bm x'\!, \!\bm x)dS'_k}{8\pi},\label{in}
\end{eqnarray}
where $\bm f$ is the integral over the far surface at infinity,  $S_n$ is the surface of the $n-$th particle and $dS_k$ is aligned with the outward normal to the particle surface. There is no contribution from the channel boundaries $z=0$ and $z=h$ since both flows vanish there.

We assume that the particles are confined in a finite region so that the flow far from the particles is the Poiseuille flow given by Eq.~(\ref{ps}), see Eq.~(\ref{ds}). Since the Stokeslet decays far from the source, the second and the third terms on the RHS of Eq.~(\ref{in}) decay to zero at large distances from the particles.
 Thus, the asymptotic approach of $\bm u$ to the Poiseuille flow at large distances implies the asymptotic equality of $\bm f$ to the Poiseuille flow. It is readily seen using the asymptotic form of the Stokeslet at large horizontal distances, provided in the next Section, and the asymptotic, Poiseuille, form of the flow, that $\bm f$ is determined by terms independent of the particles, that is terms that would be the same for the case of no particles.
Thus necessarily $\bm f$ is the Poiseuille flow given by Eq.~(\ref{ps}). This is confirmed by the direct calculation in Appendix \ref{f}. We therefore conclude that,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
u_l\!=\!\frac{\delta_{lx}z(z\!-\!h)\nabla_x p^0}{2\eta}\!-\!\sum_n\int_{S_n}\!\!\!\frac{S_{il}(\bm x', \bm x)\sigma'_{ik}dS_k'}{8\pi \eta}\nonumber\\&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
+\sum_n\int_{S_n}\frac{u_{i}(\bm x')T_{ilk}(\bm x'\!, \!\bm x)dS'_k}{8\pi}.\label{inrep}
\end{eqnarray}
This integral representation of the flow involves no approximations and holds for particles of arbitrary shape. The flow is determined by the values of $\bm u$ and $\sigma_{ik}$ at the particle surface. When the distance between the particles are much larger than their size the hydrodynamic interactions are negligible and $\bm u$ and $\sigma_{ik}$ are approximately those of an isolated particle. However, the current representation works also when the particles are close to each other so that the hydrodynamic interactions change significantly the values of $\bm u$ and $\sigma_{ik}$ at the particle surface.

%We can rewrite the representation given by Eq.~(\ref{inrep}) using the form of $T_{ilk}$ given by Eq.~(\ref{strt}). The term involving the pressure is
%\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
%-\int_{S_n}\frac{p_l(\bm x'\!, \!\bm x)\bm u(\bm x') \cdot d\bm S'}{8\pi}=\int \frac{dV'}{8\pi}\bm u(\bm x')\cdot \nabla' p_l(\bm x'\!, \!\bm x).
%\end{eqnarray}

{\it Simplification for rigid surface.}---The integral representation above simplifies in the case of rigid surfaces when the flow on the surface of the particles is the superposition of translation and solid-body rotation. This is not only the case of rigid particles described after Eq.~(\ref{ds}), but  often also the case of small bubbles where impurities present in the fluid accumulate at the surface making it effectively rigid. In this case, experiments demonstrate that the behavior of these bubbles is similar to that of rigid particles \cite{hb,levich}. In these and similar cases the flow at the particle surface $u_i(\bm x')$ has the form  $c_i+\epsilon_{irm}x'_m{\tilde c}_r$ where both $\bm c$ and $\bm {\tilde c}$ are independent of $\bm x'$. Hence the last term in Eq.~(\ref{inrep}) drops out because  for $\bm x$ outside the particle interior $V_p$,
\begin{eqnarray}&&
\int_{S_p}dS'_k T_{ilk}(\bm x', \bm x)=\int_{V_p} dV \nabla'_k T_{ilk}=0, \\&& \epsilon_{irm}\int_{S}dS'_k x'_mT_{ilk}(\bm x', \bm x)
=\epsilon_{irm}\int_{V_p} dV \delta_{mk}T_{ilk}=0,
\nonumber
\end{eqnarray}
cf.\ \cite{ps}. Note that we made use of the fact that $\epsilon_{irm}$ is antisymmetric over indices $i$, $m$ whereas $T_{ilm}$ is symmetric with respect to those indices. We conclude that in the case of many particles with rigid surfaces,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
u_l\!=\!\frac{\delta_{lx}z(z\!-\!h)\nabla_x p^0}{2\eta}\!-\!\sum_n\int_{S_n}\!\!\!\frac{S_{il}(\bm x', \bm x)\sigma'_{ik}dS_k'}{8\pi \eta},\label{inrepmany}
\end{eqnarray}
The representations derived here are thus useful to descrive  the  flow.

\section{Derivation of lubrication theory}

In this Section we demonstrate that Eqs.~(\ref{inrep}) and (\ref{inrepmany}) imply that the predictions of the lubrication theory hold at horizontal distances from the particles larger than $h$. This is less restrictive than the usual condition of applicability of the lubrication theory for distances much larger than $h$, cf.\ \cite{is}. This property comes from the Stokeslet flow which obeys the lubrication theory at distances larger than $h$. We use the representation
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!
S_{il}(\bm x', \bm x)\!=\!\frac{12 z'(h\!-\!z')z(z\!-\!h)\nabla_i\nabla_l  \ln \rho}{h^3}\!+\!{\tilde S}_{il}(\bm x', \bm x),\label{Stokesl}
\end{eqnarray}
where $\rho$ is the length of $\bm \rho\!=\!(x-x', y-y')$ (thus $\nabla_i\nabla_l  \ln \rho=0$ if one of the indices is $z$). It was observed in \cite{LironMochon}
that ${\tilde S}_{il}$ decays exponentially in $\rho$ with exponent at least $\pi/h$, that is the smallness is at least $\exp(-\pi\rho/h)$.
%Similarly, if one of the indices of $S_{il}$ is $z$ then $S_{il}$ is of order $\exp(-\pi\rho/h)$.
Thus already at $\rho\approx h$ we can discard the last, non-potential term in Eq.~(\ref{Stokesl}). The resulting approximation to the Stokeslet flow,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!
S_{il}(\bm x', \bm x)\!\approx\! S^0_{il}(\bm x', \bm x)\!= \!\frac{12 z'(h\!-\!z')z(z\!-\!h)\nabla_i\nabla_l  \ln \rho}{h^3},\label{so}%\nonumber
\end{eqnarray}
is the two-dimensional potential flow times $z'(h-z')$, in agreement with the predictions of the lubrication theory, see Eq.~(\ref{lubr}). Note  that $S^0_{il}(\bm x', \bm x)$ is a symmetric function of $\bm x$ and
$\bm x'$, whose dependence on the horizontal coordinates is via the difference $\bm \rho$ only. We consider the corresponding pressure $p^0$ that approximately solves the corresponding Stokes equation
$\nabla'_i p^0_l(\bm x', \bm x)\!=\!\nabla'^2 S^0_{il}(\bm x', \bm x)$, see the second of Eqs.~(\ref{st}). Here $\nabla'_i$ designates the derivative over $x'_i$ and we do not write the $\delta(\bm x'-\bm x)$ term on the RHS. We thus find ($l\neq z$),
\begin{eqnarray}&&
p^0_l(\bm x', \bm x)=\frac{24 z(z-h)\rho_l}{h^3 \rho^2}=\frac{24z(z-h)}{ h^3}\nabla_l\ln \rho.
\end{eqnarray}
%as can be checked readily.
%We check this,
%\begin{eqnarray}&&
%\nabla'_i p^0_l(\bm x', \bm x)=-\frac{24z(z-h)}{ h^3}\nabla_i\nabla_l\ln \rho,\\&&
%\nabla'^2 S^0_{il}(\bm x', \bm x)=-\frac{24z(z\!-\!h)\nabla_i\nabla_l  \ln \rho}{h^3}.
%\end{eqnarray}
Here, $p^0$ is the leading order approximation for the pressure of the Stokeslet at large distances \cite{LironMochon}, with an exponentially small correction.
We can write ($l\neq z$ but $i$ or $k$ can be  $z$),
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!
T_{ilk}(\bm x', \bm x)=-p^0_l\delta_{ik}+\nabla'_k S^0_{il}+\nabla'_i S^0_{kl}\!+\!{\tilde T}_{ilk}(\bm x', \bm x),
\end{eqnarray}
where ${\tilde T}_{ilk}(\bm x', \bm x)$ decays exponentially in $\rho$ with exponent at least $\pi/h$, cf.\ Eq.~(\ref{st}). The stress tensor $T_{ilk}(\bm x', \bm x)$
is exponentially small when one of the indices is $z$.
We find using the expressions for $p_0$ and $S^0_{il}$,
\begin{eqnarray}&&
T_{ilk}(\bm x', \bm x)\approx \frac{24z(h-z)}{h^3}\nabla_l\left[ \left(\delta_{ik}+z'(h\!-\!z')\nabla_i \nabla_k \right)\ln \rho\right]\nonumber\\&&
+\frac{12 (h\!-\!2z')z(z\!-\!h)}{h^3}\left(\delta_{kz}\nabla_i+\delta_{iz}\nabla_k \right) \nabla_l\ln \rho,
\end{eqnarray}
where we neglected exponentially small correction. We find from Eq.~(\ref{inrep}) that,
\begin{eqnarray}&&\!\!\!\!\!\!\!\!\!\!
\bm u(\bm x)\!=\!\frac{z(z\!-\!h)\nabla p}{2\eta}+O\left(e^{-\pi\min[\rho_n]/h}\right)\!;\ \ \nabla^2 p\!=\!0,\label{fi}
\end{eqnarray}
where  
$\bm u=(u_x, u_y)$ and $\min[\rho_n]$ is the distance from $\bm x$ to the closest boundary of a particle. 
The pressure $p$ in this formula is independent of $z$ as predicted by the lubrication theory used in Eq.~(\ref{lubr}) with,
\begin{eqnarray}&&\!\!\!\!\!\!
p=p^0+\sum_n \delta p_n,\ \ \delta p_n=\frac{3\nabla_i}{\pi h^3}\int_{S_n}z'(z'-h) \ln \rho \sigma_{ik}'dS_k'\nonumber
\\&&\!\!\!\!\!\!
-\frac{6\eta}{\pi h^3} \int_{S_n} dS'_k  \left(\delta_{ik}+z'(h\!-\!z')\nabla_i \nabla_k \right) u_{i}(\bm x') \ln \rho\nonumber\\&&\!\!\!\!\!\!
+\frac{3\eta}{\pi h^3} \int_{S_n} dS'_k  (h\!-\!2z')\left(\delta_{kz}\nabla_i+\delta_{iz}\nabla_k \right) u_{i}(\bm x') \ln \rho,\label{fu}
\end{eqnarray}
where $\delta p_n$ is the pressure perturbation due to the $n-$th particle and the summation over repeated indices is from $1$ to $3$. Eqs.~(\ref{fi})-(\ref{fu}) are one of the main results of our work. These provide a refinement of the lubrication theory demonstrating that Eq.~(\ref{lubr}) holds under
the condition $\exp{(-\pi \min[\rho_n]/h)}\ll 1$, which is difficult to show
using the classic lubrication theory as it demands the strong inequality $\min[\rho_n]\gg h$. For instance, at $\min[\rho_n]=h$ the exponential factor is $\sim 0.04$. The result holds both for droplets and rigid particles where for rigid particles the last two lines of Eq.~(\ref{fu}) become zero and the equation reduces to
\begin{eqnarray}&&\!\!\!\!\!\!
p=p^0+\frac{3\nabla_i}{\pi h^3}\sum_n\int_{S_n}z'(z'-h) \ln \rho \sigma_{ik}'dS_k'.\label{rg}
\end{eqnarray}
We have good control of the correction terms to Eqs.~(\ref{fi})-(\ref{rg}) from the series representation of ${\tilde S}_{il}(\bm x', \bm x)$ provided in \cite{LironMochon}.

The pressure $p$ solves the two-dimensional Laplace equation in the domain between the particles since it is formed by integrals of the fundamental solution of the Laplace equation $\ln \rho$ over the particle boundaries.
The formula for $p$ matches the ideal flow that holds beyond the horizontal distance $h$ from the particles with the fully viscous flow near the particles. The viscous layer is the neighborhood
of the boundary of each particle where Eq.~(\ref{fu}) breaks down. Though the solution for $p$ is given in terms
of the unknown velocities and stress tensors on the surfaces of the particles, it seems that this is as much as can be done generally: the matching problem is not solvable for any general particle shape. It does simplify for disk-like particles as described previously.

{\it Hydrodynamic interactions of pancake-like droplets.}---Eqs.~(\ref{fi})-(\ref{fu}) provide support for the observation that  the width of the viscous layer around disk-like droplets is not larger than $h$.
The formulae tell that, unless the distance between the droplet surfaces is smaller than $h$, the (horizontal) flow outside the viscous layers near the particles is an
ideal potential flow. This flow can be determined using the boundary condition that the normal velocity at the outer boundary of the viscous layer coincides with the normal component of the translational velocity of the particle.
%%
Since the layer width is of the order of $h$, and as long as the distance between the droplets is larger than $h$ (but possibly much smaller than $a$) we can impose the boundary condition on the particle surface, neglecting the finite width of this viscous layer as we did for the case of the single large disk, see Eq.~(\ref{dl}). Similarly, in the presence of many particles whose separation is larger than $h$, the flow outside the boundary layer is described by a pressure field $p$ that obeys \cite{is,tl2012},
\begin{eqnarray}&&%\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!
\nabla^2 p=0,\ \ \left(\bm v_n+\frac{h^2\nabla p}{12\eta}\right)\cdot{\hat n}_n=0, \label{press}
\end{eqnarray}
where $\bm v_n$ is the velocity of $n-$th particle, ${\hat n}_n$ is the unit vector normal to the surface of the $n-$th particle. The pressure gradient is taken at the outer boundary of the viscous layer of the $n-$th particle. However, since the latter is narrow, one can consider $\nabla p$ on the surface of the $n-$th particle without affecting significantly the solution for the pressure outside the viscous layers. To find the pressure inside the layers would require a separate study. For close droplets the pressure determined by Eq.~(\ref{press}) is different from the superposition of the dipole solutions given by Eqs.~(\ref{dlp0})-(\ref{dl}) due to the near-field interactions.

>>>>

%--%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%--%%This is the LaTeX ARTICLE template for RSC journals
%--%%Copyright The Royal Society of Chemistry 2016
%--%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
